---
title: "ACM School Placement"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install Packages

```{r install packages, echo = FALSE}
# install.packages("gmapsdistance")
# install.packages("readxl")
# install.packages("googleway")
# install.packages("dplyr")
# install.packages("tidyr")
# Nick mentioned there may be a useful package that breaks down script timings 
```

# Load Packages

```{r load packages}
library("gmapsdistance")
library("readxl")
#library("googleway")
library("dplyr")
library("tidyr")
```

# Read Data

This is fake data that may or may not represent an actual corps.

```{r load spreadsheets}
acm_df <- read_excel(path = "Input 1 - ACM Data.xls")
acm_commutes <- read_excel(path = "Input 2 - ACM Commutes.xlsx")
school_df <- read_excel(path = "Input 3 - School Data.xls")

# For some reason there is an empty row at the end of acm_df, so it's removed here:
acm_df <- acm_df[-nrow(acm_df),]

# Add id columns
acm_df$id <- 1:nrow(acm_df)
acm_commutes$id <- 1:nrow(acm_commutes)
school_df$id <- 1:nrow(school_df)

```

# Calculate Commutes

To calculate commute times, we essentially send many requests to Google Maps through an API (Applciation Program Interface). To get access to the API we need a key. Visit the [Google Distance Matrix API](https://developers.google.com/maps/documentation/distance-matrix/) page. Click "GET A KEY" in the upper right corner. Save the key privately. This key identifies you to Google.

Many organizations provide access to their data through an API. Many times this service is free, but some organizations charge for this service because of processing costs and the benefits that this access enables. Google will allow you to calculate 2,500 travel times for free in one day. After that, you will need to pay $0.50 for each 1,000 requests, or wait a day to calculate 2,500 more.

The key I included in this script is a free-use key, so it will be limited to 2,500 results per day. However, I did opt to enable pay-as-you-go on my personal key (not shared here). In CYCHI, this expense will hopefully be covered under our Impact Team budget.

```{r commute times}
# Prior to this step, you may need to ensure that "Address Line 1" 
# does not include any apartment/suite/room info.

# Note that for public tansit, google does not count 
# the time spent waiting for the bus/train. For example, 
# let's say I get off work at 5pm and a bus comes at 5:15pm. 
# I ride the bus 45 minutes to get home. My commute is more 
# accurately described as 60 minutes, but this function would 
# still return 45 minutes.

run_commutes = FALSE

if (run_commutes == TRUE){

  # This line combines address data into one text string:
  acm_df$"Full Address" = paste(acm_df$"Address Line 1", acm_df$"City", acm_df$"State", acm_df$"Postal Code")
  
  # Replace spaces with "+" and remove commas (requests to google maps API cannot include spaces)
  acm_df$"Full Address" = gsub(" ", "+", acm_df$"Full Address")
  school_df$"Address" = gsub(" ", "+", school_df$"Address")
  school_df$"Address" = gsub(",", "", school_df$"Address")
  
  # Input your own API Key here
  set.api.key("AIzaSyDFlU9RkmJBJdw0YGMswYECXQeZeKxFmuc")
  
  # Create an empty dataframe that we will fill with commute times
  acm_commutes <- data.frame()
  
  # Create a for loop that will read through each row of ACM data, feed it into the main function of our gmapsdistance package, and build a new data frame of commute info.
  for (acm_id in acm_df[1:2,]$id){
    
    # select just one row from acm_df, and assign it to a new object, acm_row
    acm_row <- subset(acm_df, id == acm_id)
    
    # feed that ACM's address and mode into the function 'gmapsdistance'. This will return a new object that is a single row of ACM commute times to each school. That row is assigned to a new object, 'commute'
    commute = gmapsdistance(origin = acm_row$"Full Address", destination = school_df[1:2,]$"Address", mode = acm_row$"Method of Commute", combinations = "all", shape = "wide")
    
    # create an 'id' column in our new 'commute' row that is the same as acm_id. We will use this to join our data frames.
    commute$Time[["id"]] = acm_id
    
    # as the for-loop runs, progressively add each single row of commute data into a new data frame called acm_commutes. As this for-loop runs, this dataframe grows to include all ACM's.
    acm_commutes <- rbind(acm_commutes, commute$Time)
  }
}

```

# Initial Team Placements
  
Generate a list of team id numbers for each team that is the length of each team's size. Combine these lists into one long list of initial team assignments.
  
```{r}  
# Assumption: the number of ACMs in acm_df is exactly the same as the number of team slots

# first create an empty list
team_placements = list()

# use a for-loop to read each team size
for (x in 1:nrow(school_df)){
  team_slots = list(
    # create a list that repeats each school 'id' for the size of each team
    rep(x, 
        subset(school_df$"Team Size", school_df$"id" == x)
        )
    )
  team_placements <- c(team_placements, team_slots)
}

team_placements <- unlist(team_placements)
# team_placements_random <- sample(length(team_placements), length(team_placements))
# team_placements_random
```

### Calculate score

This is the function that calculates the quality of each placement.

```{r}
# Chris
  # gender/sex
  # age
  # ethnicity
  # attend local school, CY or local
  # roommates? not necessary?
# Alex
  # educational attainment (potentially different goals for HS/ES)
  # tutoring experience (look at notes from survey)
  # tutoring preference
  # grade level preference
  # language speaking

PickTime <- function(myid, myplace){subset(acm_commutes,
                                              subset = id == myid,
                                              select = myplace+1)}

## We want the following characteristics to be represented evenly across teams. We calculate ratios  for the # entire corps, and later calculate how well each team matches these ratios.
# gender.ratio.ideal <- with(acm_df, sum(Gender == "Male")/sum(Gender == "Female"))
## non-binary
# caucasian.ratio.ideal <- 
# afamerican.ratio.ideal <- 
# latinx.ratio.ideal <- 

## Characteristics that we don't want to be even across all schools
# edu.attainment.hs.ideal <- 
# edu.attainment.es.ideal <- 
# spanish
  
# Commute times in long format will be easier to index and subset
acm_commutes_long <-
  select(acm_commutes, -or) %>%
  gather(dest, dist, -id) %>%
  mutate(id_dest = paste(id, dest, sep = "_"))

calculate_score = function(acm_df, team_placements) {
  
  # Convert the team_placement list into a dataframe with an 'id' column 
  team_placements_df <- data.frame(id = 1:length(team_placements),
                                   Team.Placement = team_placements,
                                   dest = colnames(acm_commutes)[team_placements + 1]) %>%
    within(id_dest <- paste(id, dest, sep = "_"))
  
  # Merge team_placements_df with acm_df on the 'id' column we just created
  acm_df_with_placements <- merge(acm_df, team_placements_df, by = "id", all.x = TRUE)
  
  # Merge acm_df_with_placements with school_df to pull in school characteristics
  acm_df_with_placements <- merge(acm_df_with_placements, school_df, by.x = "Team.Placement", by.y = "id", all.x = TRUE)
  
  #Placement.Scores <-
  #  group_by(acm_df_with_placements, Team.Placement) %>%
  #  summarize(gender.ratio = sum(Gender == "Male")/sum(Gender == "Female"),
  #            age.var = var(Age)) %>%
  #  mutate(gender.ratio.dev = abs(gender.ratio - gender.ratio.ideal)) %>%
  #  ungroup() %>%
  ## 3 is some arbitrary score weight
  #  summarize(gender.ratio.score = mean(gender.ratio.dev)*3,
  #            avg.age.var = -mean(age.var))
  ## Also, negative scores are better, and the "-" in front of mean age variance means that
  ## higher average variance of team ages gets relatively more favorable scores.
  #
  ## New commute calc method references functions defined just before 'calculate_score'
  acm_df_with_placements$"Commute.Time" <- unlist(with(acm_df_with_placements, mapply(PickTime, id, Team.Placement)))
  
  # NSM's new method for calling commute times
  system.time({
    filter(acm_commutes_long, id_dest %in% team_placements_df$id_dest) %>%
      summarize(sum(dist))
  })
 
  # NSM: even faster way to do the same thing, with the data.table package
  library(data.table)
  dt_commutes <- data.table(acm_commutes_long)
  system.time({
    dt_commutes[id_dest %in% team_placements_df$id_dest,
                sum(dist)]
  })
  
  return(sum(acm_df_with_placements$"Commute.Time"))
  
  # Previous commute calc method (10x slower!)
  #   acm_row <- subset(acm_df_with_placements, id == x)[ , grepl("Time.to.", names(acm_row))]
  #   acm_commute <- acm_row[ , grepl("Time.to.", names(acm_row))]
  #   Actual.Commute.Time <- acm_commute[acm_row$Team.Placement]
  #   acm_commutes <- rbind(acm_commutes, Actual.Commute.Time[1,1])

}

```

```{r}
current_temperature = function(iter, s_curve_amplitude, s_curve_center, s_curve_width) {
  s_curve_amplitude * s_curve(iter, s_curve_center, s_curve_width)
}

s_curve = function(x, center, width) {
  1 / (1 + exp((x - center) / width))
}

```

```{r}
run_intermediate_annealing_process = function(acm_df, team_placements, placement_score, best_placements, best_score, starting_iteration, number_of_iterations, s_curve_amplitude, s_curve_center, s_curve_width) {
  
  for(i in 1:number_of_iterations) {
    iter = starting_iteration + i
    temp = current_temperature(iter, s_curve_amplitude, s_curve_center, s_curve_width)
    
    # Create a copy of team_placements
    candidate_placements = team_placements
    
    # Create an 'index' column
    candidate_placements_df = data.frame(index=1:length(candidate_placements), candidate_placements)
    
    # Choose 2 schools at random
    schools_to_swap = sample(1:nrow(school_df), 2)
    
    # Choose 1 ACM from each of those schools
    swap1 = with(candidate_placements_df, sample(index[candidate_placements == schools_to_swap[1]], 1))
    swap2 = with(candidate_placements_df, sample(index[candidate_placements == schools_to_swap[2]], 1))
    # Swap the team assignment of those 2 ACMs
    candidate_placements = replace(candidate_placements, c(swap1, swap2), candidate_placements[c(swap2, swap1)])
    
    # Merge in ACM data
    candidate_score = calculate_score(acm_df, candidate_placements)

    if (temp > 0) {
      ratio = exp((placement_score - candidate_score) / temp)
    } else {
      ratio = as.numeric(candidate_score < placement_score)
    }
    
    if (runif(1) < ratio) {
      team_placements = candidate_placements
      placement_score = candidate_score
      
      if (placement_score < best_score) {
        best_placements = team_placements
        best_score = placement_score
      }
    }
  }
  
  return(list(team_placements=team_placements, placement_score=placement_score, best_placements=best_placements, best_score=best_score))
}

```

```{r}
## USEFULL: system.time({ <<code to time>> })

# system.time({ 
output <- run_intermediate_annealing_process(acm_df, team_placements, placement_score = 445109, best_placements = team_placements, best_score = 445109, starting_iteration = 1, number_of_iterations = 1000, s_curve_amplitude = 4000, s_curve_center = 0, s_curve_width = 3000)
#})
output
```

```{r}
# Convert the team_placement list into a dataframe with an 'id' column 
best_placements_df <- data.frame(id = 1:length(output$best_placements), Team.Placement = output$best_placements)
  
# Merge team_placements_df with acm_df on the 'id' column we just created
acm_df_output <- merge(acm_df, best_placements_df, by = "id", all.x = TRUE)
  
# Merge acm_df_with_placements with school_df to pull in school characteristics
acm_df_output <- merge(acm_df_output, school_df, by.x = "Team.Placement", by.y = "id", all.x = TRUE)

acm_df_output
```

# End Here

# Build Requirements ()

```{r generate logic for invalid assignments}
# E.g., can't allow staff with age < 21 into an HS

IsValid <- function(assign){
  valid <- ifelse(any(assign < 9), FALSE, TRUE) 
  return(valid)
}

```



```{r randomly generate initial assignment}

random.gen <- function(acm_df, school_df){
  nSch <- nrow(school_df)
  nACM <- nrow(acm_df)
  team.size <- ceiling(nACM/nSch)
  assign <- rep(school_df$id, each = team.size) # Creates a vector of 1 10 times, 2 10 times, etc
  random <- sample(x = assign, size = nStu)
    # table(random) # ... This displays a tabulation of the assigned schools
  return(random)
}

repeat{
  myAssign <- random.gen(acm_df, school_df)
  table(myAssign)
  if (IsValid(myAssign)) break
}
```

```{r generate proposal function for algorithm's steps}
GenProp <- function(myAssign){
  # picks random to and from schools, and swaps those assignments
  # returns the full assignment
}
```


```{r create objective function}
Obj <- function(myAssign){
  # Takes the assignment, and calcules an objective score combining:
  # 1. minimize commute time
  # 2. college experience or age >= 21 for HS (might also be requirement for valid proposal)
  # 3. bonus for demographic balance (e.g. gender, race/eth, ed)
  # 4. grade-level and perhaps even school preferences
  
  
}
```

```{r create a temperature schedule as a function of step number}

```

```{r implement SA algorithm}

```


* picking initial 'direction'
  * how to set up function to know that there are teams of 9 (or 8, 7, 10, etc)
  * output: series (1, 1, 1, 1, 2, 2, 2, 2, 3, 3, etc.)
* Identify right solution algorithm for our setup
  * GenSA
  * gaoptim -- is for permutation-based problems
  * roll our own (Nick has code for this)

